---
title: "class11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” 
#“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
#and determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
#you determine what proportion of structures are protein

```{r}
data<-read.csv("Data Export Summary.csv")
```
```{r}
total<-sum(data$Total)
total
```

```{r}
data$Total
```
```{r}
ans <- data$Total/sum(data$Total) * 100
names(ans) <- data$Experimental.Method
round(ans,2)
```
```{r}
round(sum(data$Proteins)/sum(data$Total) * 100, 2)
```
#working with biomolecular data in R
```{r}
library(bio3d)
```
```{r}
pdb <- read.pdb("1hsg")
```
```{r}
pdb$atom
```
```
```
```{r}
pdb$atom[1,"resid"]
```

```{r}
aa321(pdb$atom[,"resid"])
```

#select atoms Question 8. First select "protein" then write out a file: 1hsg_protein.pdb"
```{r}
#atom select()
prot<- atom.select(pdb, "protein", value=TRUE)
#write.pdb()
write.pdb(prot, file="1hsg_protein.pdb")
```
#now the same for the ligan
```{r}
#atom select(): this function  might be confusing so think about it as in vmd when we select for the ligand only or the protein. 
lig<- atom.select(pdb, "ligand", value=TRUE)
#write.pdb()
write.pdb(lig, file="1hsg_ligand.pdb")
```
```{r}
library(bio3d.view)
#view(lig)
view(pdb,"overview",col="sse")
#this package bio3d.view is not working on all platform wight now.
```
#computing connectivity from coordinates: let's try a NMA:


```{r}
# Load the package
pdb <- read.pdb("1hel")
# Normal mode analysis calculation
modes <- nma(pdb)
m7 <- mktrj(modes,
 mode=7,
 file="mode_7.pdb")
view(m7, col=vec2color( rmsf(m7) ))
```







