---
title: "unsupervisedproject"
author: "Giulia Livrizzi"
date: "10/30/2019"
output: html_document
---
##Data input
```{r}
library(readr)
WisconsinCancer <- read.csv("WisconsinCancer.csv")
View(WisconsinCancer)
```
```{r}
wisc.df<-read.csv("WisconsinCancer.csv")
head(wisc.df)
View(wisc.df)
```

here we examine data from `r nrow(wisc.df)` patient samples
```{r}
knitr::opts_chunk$set(echo=TRUE)
```
```{r}
class(read.csv("WisconsinCancer.csv"))
```

```{r}
table(wisc.df$diagnosis)
```

In this data-set we have `rx["M"]` cancer and `r["B"]` non-cancer

```{r}
wisc.data<-as.matrix(wisc.df[,3:32])
```
```{r}
#access only mean columns
colnames(wisc.df)
```
too look for patterns (like which column has a mean value) you can call `grep()`
```{r}
grep("_mean", colnames(wisc.df), value = TRUE)
```
TO FIND OUT HOW MANY THERE ARE I CAN CALL `lenght()` on the result of `grep()`
```{r}
lenght (grep("_mean", colnames(wisc.df), value = TRUE))
```
performing PCA. Do we need to scale the data? this runs apply over the 2 columns and round the 3
```{r}
round(apply(wisc.data,2,sd),3)
```
looks like we need to `scale=TRUE` since our data are like everywhere(i.e sd values were too different from each other)
```{r}
wisc.pr<-prcomp(wisc.data, scale=TRUE)
summary(wisc.pr)
```
Plot PC1 vs PC2 and color by M/B cancer/noncancer diagnosis with no colors
```{r}
plot (wisc.pr$x[,1], wisc.pr$x[,2])
```
Plot PC1 vs PC2 and color by M/B cancer/noncancer diagnosis with  colors
```{r}
plot(wisc.pr$x[,1:2], col=wisc.df$diagnosis)
```
```{r}
x<-summary(wisc.pr)
```
the first PC captures `x$importance[,"PC1"]*100` of the original variance in the dataset
```{r}
x$importance[,"PC1"]

```
Q8. How many principal components (PCs) are required to describe at least 70% of the original variance in the data? here I am looking for where my numbers get over 0.70. this cose will return logicals like TRUE or FALSE.
```{r}
which(x$importance[3,]>0.7)
```
we can score genes based on how much they influence PC1. Cell1 PC1 score = (read count * influence) + … for all genes

```{r}
data.scaled <-scale(wisc.data)
```
```{r}
data.dist <- dist(data.scaled)
```
```{r}
hclust(dist(wisc.pr$x[,1:3]))
y<-hclust(dist(wisc.pr$x[,1:3]))
plot(y)
```
```{r}
data.scaled<-scale(wisc.data)
wisc.hclust<-hclust(dist(data.scaled))
plot(wisc.hclust)
```
```{r}
grps <- cutree(y, k=2)
table(grps)
```
let's try clustering in PCA space since this looks bad
we will take the results of `prcomp()` and build our distance matrix in PC space rather than from our raw data
take the first 7 PCs
```{r}
wisc.pr.hclust<-hclust(dist(wisc.pr$x[,1:7], method="ward.D2"))
plot(wisc.pr.hclust)
```
and the cluster member vector might be obtained from `cutree()`
```{r}
grps<-cutree(wisc.pr.hclust,h=70)
table(grps)
```
```{r}
table(d=wisc.df$diagnosis)

```
let's compare i.e. cross tabilate these results
```{r}
table(grps,wisc.df$diagnosis)
```



