---
title: "machine learning"
author: "Giulia Livrizzi"
date: "10/25/2019"
output: html_document
---
##K-mean examples. Baby steps for machine learning (make up some data to cluster)
```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3)) #data are going to be flipped around -3/3
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```
#Use the kmeans() function setting k to 2 and nstart=20
#Inspect/print the results
#Q. How many points are in each cluster?
#Q. What ‘component’ of your result object details
# cluster size?
#cluster assignment/membership?
#- cluster center?
#Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
k<-kmeans(x,centers=2,nstart=20)
```
```{r}
k$cluster
```
```{r}
table(k$cluster)
```
```{r}
plot(x,col=k$cluster)
##funciont point will add points to an existing plot
points(k$centers, col="blue",pch=15)

```


#LET'S MOVE ON TO HIERARCHICAL CLUSTERING
```{r}
hc <-hclust ( dist(x))
```
#plot my results
```{r}
plot (hc)
abline(h=6, col="red")
abline(h=4, col="blue")
cutree(hc,h=6)
```
```{r}
grps<-cutree(hc,h=4)
table(grps)
```
#I can also cut the tree to yeld a given 'k' groups/clusters
```{r}
cutree(hc,k=2)
```
```{r}
plot(x, col=grps)
```


```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
```

```{r}
# Step 2. Plot the data without clustering
plot(x)
```

```{r}
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)
```
```{r}
hc <-hclust ( dist(x))
```
#plot my results
```{r}
plot (hc)

```







