---
title: "class14"
author: "Giulia Livrizzi"
date: "11/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
read.csv("airway_metadata.csv")
```

```{r}
read.csv("airway_scaledcounts.csv")
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```
#SETUP Bioconductor DESeq2
```{r eval=FALSE}
install.packages("BiocManager")
BiocManager::install()
BiocManager::install("DESeq2")
```
```{r}
head(metadata)
```
how many genes do we have in this data set?

```{r}
nrow(counts)
```
how many experiments?
```{r}
ncol(counts)-1
```
let's make sure metadata id col matches the colnames of counts
```{r}
colnames(counts)[-1]
```
```{r}
metadata$id
```

```{r}
all(colnames(counts)[-1] ==metadata$id)
```
all function to look across a vector of logicals
```{r}
all(c(T,T,T))
```
#Analysis: compare averages of controls are different from treated averages. First we need to access the columns of our countdata that are control and then find their mean.
```{r}
metadata$dex =="control"
```
#these are the columns I want to access
```{r}
control.id <-metadata[metadata$dex =="control", ]$id
control.id
```

```{r}
head(counts[,control.id])
```
#mean counts for control experiments(lenght(control.id) is number of columns -i.e. 4 here)
```{r}
control.mean <-rowSums(counts[,control.id])/length(control.id)
names(control.mean) <- counts$ensgene
```
#Do the same for drug treated exps
```{r}
treated <- metadata[metadata [,"dex"]=="treated",]
treated.mean <- rowSums(counts[, treated$id])/4
names(treated.mean) <- counts$ensgene
```
```{r}
meancounts <-data.frame(control.mean, treated.mean)
```
```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```
#Good case for a log plot
```{r}
plot(meancounts$control.mean, meancounts$treated.mean, log="xy")
```
#add new column called log2fc to see the fold change
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
#We want to remove NaN and the zero points from further analysis. Which will return only the values that you are looking for.
#test for finding zero entries
```{r}
x<- c(5,8,0,5)
x==0
which(x==0)
```
```{r}
x<-data.frame(happy=c(5,6,0,0), sad=c(0,5,5,0))
x==0
```
#arr.ind gives you more insights on the position (row and column) of the NAN or zero values. you can use it even to just know the position of a gene. unique allows you to exclude those
```{r}
unique(which(x==0, arr.ind = TRUE)[,1])
```
#focus on rows that only have zero entries
```{r}
inds<-unique(which(x==0, arr.ind = TRUE)[,1])
x[-inds,]
```
#write another test
```{r}
x<-data.frame(analine=c(1,2,0,0), giulia=c(0,0,3,0))
x==0
```
```{r}

which(x==0, arr.ind = TRUE)
```
```{r}
unique(which(x==0, arr.ind = TRUE)[,1])
```

# now apply it to our metacounts data (i.e. removing NaNs and zero)
```{r}
to.rm<-unique(which(meancounts[,1:2]==0, arr.ind=TRUE)[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
#How many genes do I have left?
```{r}
nrow(mycounts)
```
#A common threshold used for calling something differentially expressed is a log2(FoldChange) of greater than 2 or less than -2. Let’s filter the dataset both ways to see how many genes are up or down-regulated.
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```
#To count how many are left:
```{r}
sum(up.ind)
sum(down.ind)
```
#In cases where you don't have a preferred annotation file at hand you can use other Bioconductor packages for annotation.
#Bioconductor's annotation packages help with mapping various ID schemes to each other. Here we load the AnnotationDbi package and the annotation package org.Hs.eg.db.

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```
```{r}
columns(org.Hs.eg.db)
```
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts), # Our gene names
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```
```{r}
head(mycounts)
```
#We still need to work on statistical tests. DESeq packages does everything we did for us included statistics. But in this way we learned a lot...

```{r}
library(DESeq2)
```
#We will use the DESeqDataSetFromMatrix() function to build the required DESeqDataSet object and call it dds, short for our DESeqDataSet. 
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)

```
```{r}
dds <- DESeq(dds)
```
```{r}
res<-results(dds)
res
```
#let's make a vulcano plot: summary figures are frequently used to highlight the proportion of genes that are both significantly regulated and display a high fold change.Typically these plots shows the log fold change on the X-axis, and the −log10 of the p-value on the Y-axis (the more significant the p-value, the larger the −log10 of that value will be).
```{r}
plot(res$log2FoldChange, res$padj)
```
#the previous plot is not good because we are not seeing the data under p<0.05 which are the significant ones. Changing the axis would not work...we have to flip it adding -log
```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
```
#We want color coding for significan changes and thresholds for -2 and 2 as well in the graph
```{r}
plot(res$log2FoldChange,  -log(res$padj), 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")
# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)
```
```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

# Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

# Cut-off lines
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```

