---
title: "GiuliaLivrizziHomework"
author: "Giulia"
date: "10/18/2019"
output: github_document
---
#WORKING ON QUESTION 6 FOR HOMEWORK
```{r}
library(bio3d)
```
```{r}
#this will be my input that I will apply:
#pdbid <- "4AKE" (I can change this depending on the kinase)
#This function will make possible to visualize B factor trends for different proteins. 
myfun <- function(pdbid, chain="A") {
  s1 <- read.pdb(pdbid) # pdbid of any protein of interest; chain A is already defined
  s1.chainA <- trim.pdb(s1, chain=chain, elety="CA")
  s1.b <- s1.chainA$atom$b #defining atoms
  plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor") #plotting the results.
}
```
```{r} 
myfun("1AKE") #testing if the function works again with another kinase name
```


#Down here working on the other questions in class (not for homework):



```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```
#Q1:when we run read.pdb(nameofprotein) it  searches online for the PDB file and it imports the data
#Q2: when we use trim.pdb,we can obtain a smaller subset of atoms from the large protein. In the argument you have to specify the chains and the targeted atoms and specify the element type(elety=CA). 
#Q3:Troubloshooting:removing the black and grey boxes....but what are they?secondary structures forms. 
#Q4:put them on the same graph
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s3.b, typ="l", col="blue")
points(s2.b, typ="l",  col="red")
#Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? I corrected the mistake in the code to have s1,s2,s3 and so the more similar ones will be 1AKE and 1E4Y. 
#create a function that plot them to quantify:
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)


