---
title: "class07"
author: "Giulia Livrizzi"
date: "10/23/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
source("http://tinyurl.com/rescale-R")
```

```{r}
rescale(1:10)
```

```{r}
rescale(c(3,10, NA,7, "giulia"))
```
#understand the stop function which is built in our code. It allows you to give a more specific comment if an error occurs. 
```{r}
rescale2 (c(3,10, NA,7, "giulia"))
```
# "!" will flip something in true/false
#We want to write a function, called both_na(), that counts how many positions in two input vectors, x and y, both have a missing value. start with a simple definition of x and y. You can use & (means and) to check at the same time 2 variables.

```{r}
x<- c (1,2,NA, 3,NA)
y<-c( NA,3,NA,3,4)

is.na(x) & is.na(y)
```
```{r}
which(is.na(x))
```
#let's start making the code more straightforward
```{r}
both_na<- function(x,y) {
  sum(is.na(x)$is.na(y))
}
```
#number of cases where we have both NA but we have to check the lenghts before to make them match
```{r}
both_na2 <- function(x, y) {
 if(length(x) != length(y)) {
 stop("inputs do not have the same lenght")
 }
 sum( is.na(x) & is.na(y) )
}
```

```{r}
# student 1
s1<-c(100, 100, 100, 100, 100, 100, 100, 90)
# student 2
s2<-c(100, NA, 90, 90, 90, 90, 97, 80)
which.min (s1)
#exclude it from the average calculations
mean(s2[-which.min(s2)],na.rm=TRUE)
```
```{r}
x<- s2
sum(x[-which.min(x)], na.rm=TRUE) #TOTAL MINUS THE WORST
sum(x[-which.min(x)], na.rm=TRUE/lenght(x)-1)

```


